---
title: "Desafio Solarview"
author: "Semar Augusto"
date: "10/22/2018"
output:
  pdf_document: default
  html_document: default
---

Sobre o banco de dados, o arquivo pjm_est j?? possui uma tabela com um banco de dados que pode ser fornecido como uma tabela. Apenas alguns ajustes precisam ser feitos, como: pjm_load ?? uma coluna que corresponde ?? soma da venda de todas as outras regi??es, ent??o ?? interessante que o SQL fa??a esse trabalho sozinho e n??o dependa que o usu??rio do banco fa??a esse trabalho manualmente todas as vezes.

# Loading libraries and reading data
```{r, echo=FALSE, include=FALSE}
require(DT)
require(fpp2)
require(data.table)
require(tidyverse)
require(prophet)
require(forecast)

pjm_est <- read.table("data/pjm_hourly_est.csv", sep=",", header=T)
pjm_load <- read.table("data/PJM_Load_hourly.csv", sep=",", header=T)
pjme <- read.table("data/PJME_hourly.csv", sep=",", header=T)
pjmw <- read.table("data/PJMW_hourly.csv", sep=",", header=T)

pjm_est$datetime <- pjm_est$Datetime
pjm_est <- separate(pjm_est, datetime, into=c("Date", "Hour"), sep = " ")

pjm_load$datetime <- pjm_load$Datetime
pjm_load <- separate(pjm_load, datetime, into=c("Date", "Hour"), sep = " ")

pjme$datetime <- pjme$Datetime
pjme <- separate(pjme, datetime, into=c("Date", "Hour"), sep = " ")

pjmw$datetime <- pjmw$Datetime
pjmw <- separate(pjmw, datetime, into=c("Date", "Hour"), sep = " ")

pjm_est$sum <- rowSums(pjm_est[,2:13], na.rm=T)
pjm_est <- pjm_est[with(pjm_est, order(Date)),]
```

```{r}
head(pjm_est)
summary(pjm_est)
```

# Creating time related features and visualizing

## Day of the week visualization
```{r}
pjm_est$dow <- weekdays(as.Date(pjm_est$Datetime))
ggplot(pjm_est, aes(x=dow, y=sum)) + geom_boxplot()
```

## Day of the month visualization
```{r}
pjm_est$dom <- strftime(as.Date(pjm_est$Date), format= "%d")
ggplot(pjm_est, aes(x=dom, y=sum)) + geom_boxplot()
```

We can see that the fluctuations during the week is possibly what generates a fluctuation during the month.

## Full month visualization
```{r}
pjm_est$month <- strftime(as.Date(pjm_est$Date), format="%m")
ggplot(pjm_est, aes(x=month, y=sum)) + geom_boxplot()
```

## Hour of the day visualization
```{r}
ggplot(pjm_est, aes(x=Hour, y=sum)) + geom_boxplot()
```

Work hours are the ones that consume the most energy, as was expected


Lets check if the the behaviour during winter is the same as during summer.

# Winter vs Summer
```{r}
verao <- subset(pjm_est, month=='07')
inverno <- subset(pjm_est, month=='01')
nrow(verao)
nrow(inverno)

ggplot(verao, aes(x=Hour, y=sum)) + geom_boxplot()
ggplot(inverno, aes(x=Hour, y=sum)) + geom_boxplot()
```

During winter the energy consumption is more stable during the day, while in the summer there are more peaks

```{r}
ggplot(verao, aes(x=dom, y=sum)) + geom_boxplot()
ggplot(inverno, aes(x=dom, y=sum)) + geom_boxplot()
```

Month behaviour is approximately the same.

```{r}
ggplot(inverno, aes(x=dow, y=sum)) + geom_boxplot()
ggplot(verao, aes(x=dow, y=sum)) + geom_boxplot()
```

Week behaviour is also similar

Summer vacations seem to not change much the energy consumption.

## Temporal series
```{r}
train_sample <- setDT(pjm_est)
time <- ts(train_sample[,"sum"], start=1998,frequency=365*24)
autoplot(time)
```

Too much variation until 2005. The company grew a lot and that may bias the predictor. Lets get a subset of the data.

```{r}
pjm_est$Date <- as.Date(pjm_est$Date)
df <- subset(pjm_est, Date > as.Date("2005-01-01"))
new_tra <- setDT(df)
new_time <- ts(new_tra[,"sum"], start=2002, frequency=365*24)
autoplot(new_time)
```

Other good alternative would be to get a smaller region like PJM East.
```{r}
pjme <- pjm_est[,c(1,11,14,15,16,17,18,19)]
pjme <- pjme[complete.cases(pjme),]
tra <- setDT(pjme)
time <- ts(tra[,"PJME"], start=2002, end=2018, frequency=365*24)
autoplot(time)
```
However, we can see that PJME's behaviour is much different from the company as a whole.

```{r}
summary(pjm_est$PJME/pjm_est$sum)
nrow(pjme)/nrow(pjm_est)
nrow(df)/nrow(pjm_est)
```

A more complex analysis could make a predictor for every region by itself, or considering that hollidays may change consumption. However, in this notebook I will make a simple predictor without any of this hypotesis.

# Predictive Analysis
```{r}
splitting_date <- as.Date('2014-06-01')
ts_df <- ts(df[,"sum"], frequency=365*24)
train <- subset(ts_df, end=length(ts_df)-365*24*3.5)

length(train)/length(ts_df)
```

## Checking if residuals are white noise
```{r}
holt_model <- holt(ts_df)
autoplot(holt_model)
checkresiduals(holt_model)
```

p-value < 0.05, the trends are probably not white noise.

# Testing different predictive algorithms

## Naive Model

```{r}
nv <- naive(train, h=3.5*365*24)
autoplot(nv)
accuracy(nv, ts_df)
```

## ETS Model

```{r}
fcets <- ets(train)
autoplot(forecast(fcets))
accuracy(forecast(fcets), ts_df)
```

## PROPHET Model 
```{r}
splitting_date <- as.Date('2014-06-01')
train_df <- subset(df, Date < splitting_date)
stats = data.frame(y=log1p(train_df$sum), ds=train_df$Datetime)
stats = aggregate(stats$y, by=list(stats$ds), FUN=sum) # juntando se houverem 2 datetimes iguais
colnames(stats) <- c("ds", "y")
head(stats)
proph = prophet(stats)

```

### Preparing test dataset
```{r}
test_df <- subset(df, Date >= splitting_date)
test = data.frame(y=log1p(test_df$sum), ds=test_df$Datetime)
test = aggregate(test$y, by=list(test$ds), FUN=sum) # juntando se houverem 2 datetimes iguais
colnames(test) <- c("ds", "y")
```

### Making prediction
```{r}
future = make_future_dataframe(proph, periods=3.5*365*24)
prediction = predict(proph, future)
mean_pred = prediction[,c("ds", "yhat")]

prophet_plot_components(proph, prediction)
```

### Retransforming data to MW
```{r}
mean_pred$yhat = expm1(mean_pred$yhat)
head(mean_pred)
```

### Calculating accuracy.
```{r}
ts_pred <- ts(mean_pred[,"yhat"], frequency=365*24)
ts_test <- ts(test_df[, "sum"], frequency=365*24)
ts_pred <- subset(ts_pred, end=length(ts_test)) # deixando ambos com mesmo tamanho pois a predic??o foi feita com valor aproximado de per??odos
accuracy(ts_pred, ts_test)
autoplot(ts_pred)
autoplot(ts_test)
```


















